Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase with autodiff

### 2022-11-21 src/tex/ms.tex
[showyourwork] overleaf sync

\renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{relsize}
\usepackage{acro}
% \usepackage{authblk}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\newcommand\jax{\textsc{Jax}\xspace}
\newcommand\dlux{$\partial$Lux\xspace}
\newcommand\equinox{\texttt{equinox}\xspace}
\newcommand\optax{\texttt{optax}\xspace}
\newcommand\numpyro{\texttt{numpyro}\xspace}

\newcommand\kepler{\textit{Kepler}\xspace}

@@

% Describe simulation setup
% Figure 1: OPD & Pupil in our test, PSF, Flat Field histogram
Within the $\partial$Lux framework we explicitly define each operation applied to the wavefront as a unique layer. Here we employ a simple optical model with only a pupil and focal plane, defined by 6 sequential layers.

\begin{enumerate}
    \item Wavefront initialisation
    \item Application of the pupil-support
    \item Normalisation of the wavefront
    \item Application of the pupil phase mask
    \item Application of the optical aberrations
    \item Propagation to the focal plane
\end{enumerate}

@@

The wavefront then encounters low-order phase errors parameterised by coefficients on Zernike polynomials. We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms. Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \textcolor{red}{showyourwork variance}\,nm RMS total \ac{opd}. These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.

This is then propagated to the $1024\times1024$ pixel focal plane by a two-sided matrix Fourier transform \cite{Soummer2007,Martinache2020}. The detector PRF is modelled by multiplying the resulting \ac{psf} by a pixel sensitivity map which is drawn from a normal distribution with mean 0 and standard deviation 0.05. The histogram of pixel sensitivities and total pixel sensitivity map is shown in the right panel of Figure~\ref{fig:optics}.

This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths. This set of stars is then imaged four times with telescope pointing dithers applied between each image. This is done to break the natural degeneracy of stellar fluxes and pixel sensitivities. This total model requires the calculation of 240 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU. The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise. Using this set of images we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities. This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds. The bottom section shows the individual residuals for each. Clearly these parameters have been well recovered through the optimisation. Right: the recovery of the flux parameters in units of photons. The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals. \textcolor{red}{Should we address the slight overestimation of fluxes due to the PRF values being a relative measure so a small global covaraince is present?}}
    \label{fig:astro_params}
\end{figure*}

>>>>

\renewcommand{\baselinestretch}{1.0} % Change to 1.65 for double spacing
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{relsize}
\usepackage{acro}
\usepackage{listings}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

@@

\newcommand\jax{\textsc{Jax}\xspace}
\newcommand\dlux{$\partial$Lux\xspace}
\newcommand\equinox{\texttt{equinox}\xspace}
\newcommand\optax{\texttt{optax}\xspace}
\newcommand\numpyro{\texttt{numpyro}\xspace}
\newcommand\kepler{\textit{Kepler}\xspace}

@@

% Describe simulation setup
% Figure 1: OPD & Pupil in our test, PSF, Flat Field histogram
Within the $\partial$Lux framework we explicitly define each operation applied to the wavefront as a unique layer. Here we employ a simple optical model with only a pupil and focal plane, defined by 6 sequential layers. Here we show how to create this optical system within the $\partial$Lux framework:

\begin{lstlisting}[language=Python]
import jax.numpy as np
import jax.random as jr
import dLux as dl

# Define wavelengths
wavels = 1e-9 * np.linspace(545, 645, 3)

# Basic Optical Parameters
aperture = 0.5
wf_npix = 512

# Detector Parameters
det_npix = 1024
sampling_rate = 20
det_pixsize = dl.utils.get_pixel_scale(sampling_rate, wavels.mean(), aperture)

# Zernike Basis (ignore piston, tip, tilt)
zern_basis = dl.utils.zernike_basis(10, wf_npix, outside=0.)[3:]
coeffs = 2e-8 * jr.normal(jr.PRNGKey(0), (len(zern_basis),))

# Define Optical Configuration
optical_layers = [
    dl.CreateWavefront    (wf_npix, aperture),
    dl.CompoundAperture   ([aperture/2], occulter_radii=[aperture/10]),
    dl.ApplyBasisOPD      (zern_basis, coeffs),
    dl.AddOPD             (np.load("mask.npy")),
    dl.NormaliseWavefront (),
    dl.AngularMFT         (det_npix, det_pixsize)]

# Create Optics object
optics = dl.Optics(optical_layers)
\end{lstlisting}

@@

The wavefront then encounters low-order phase errors parameterised by coefficients on Zernike polynomials. We include the first 4 radial terms, ignoring global piston, tip and tilt for a total of 7 terms. Coefficients were randomly drawn from a normal distribution of mean 0 and standard deviation 20\,nm, for a total of \variable{output/opd_in.txt}\,nm RMS total \ac{opd}. These aberrations applied to the pupil mask and resulting \ac{psf} are shown in the middle panel of Figure~\ref{fig:optics}, where we see this induces a large amount of distortion in the \ac{psf}.

This is then propagated to the $1024\times1024$ pixel focal plane by a two-sided matrix Fourier transform \cite{Soummer2007,Martinache2020}. The detector PRF is modelled by multiplying the resulting \ac{psf} by a pixel sensitivity map which is drawn from a normal distribution with mean 0 and standard deviation 0.05. The histogram of pixel sensitivities and total pixel sensitivity map is shown in the right panel of Figure~\ref{fig:optics}. Here we show how to create this within the $\partial$Lux framework and create an instrument object to model this:

\begin{lstlisting}[language=Python]
# Pixel response
pix_response = 1 + 0.05*jr.normal(jr.PRNGKey(0), [det_npix, det_npix])

# Create Detector object
detector = dl.Detector([dl.ApplyPixelResponse(pix_response)])

# Combine into instrument
tel = dl.Instrument(optics=optics, detector=detector)
\end{lstlisting}

This operation is then performed over 20 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm. This set of stars is then imaged four times with telescope pointing dithers applied between each image. This is done to break the natural degeneracy of stellar fluxes and pixel sensitivities. This total model requires the calculation of 240 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU. The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and variance 2.5. This gives us an average of \textcolor{red}{showyourwork variable} photons per pixel. Using this set of images we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities. This gives us 40 stellar positional parameters (20 RA and DEC each), 20 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds. The bottom section shows the individual residuals for each. Clearly these parameters have been well recovered through the optimisation. Right: the recovery of the flux parameters in units of photons. The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals.}
    \label{fig:astro_params}
\end{figure*}
